Modular Model-Checking for Fault-Tolerant Distributed System Implementations

### 2016-11-27 doc/om-paper/paper.tex
updated omh description.

@@

To demonstrate the model and techniques, we model the classic OM(1) Byzantine fault-tolerant algorithm~\cite{} in our framework. OM(1) has been modeled in a variety of formal tools~\cite{}, providing a nice comparison with our work. Then, we make modifications to the OM(1) model, and show the model's resilience to those changes. By ``resilience'', we mean that only a small, localized portion of the must be modified, and only a small set of lemmas must be modified.

The dimensions along which we make changes are the following: the fault model, topology, local node behavior, and synchronization. \lee{Did we complete these?}

The verification is carried out in the Symbolic Analysis Laboratory~\cite{}, which contains a suite of model-checkers and a high-level specification language. In our work, we use infinite-state (SMT-based) $k$-induction~\cite{}.

The verification is interesting in its own right, as it is the first $k$-induction based verification over a fully-parameterized OM(1) model (we describe this in more detail in Section~\ref{sec:related}).

\lee{also note that OM(1) underlies many real-world systems}

\subsection{OM(1) Algorithm}

\lee{rewrite this to be OMH}

The ``Oral Messages'' algorithm solves the Byzantine Generals
Problem~\cite{Lamport-OM}. The algorithm, $\OM{m}$, is a recursive algorithm, parameterized by the number of rounds of communication, $m$. Consider a finite set of nodes $N$. Distinguish one node as the general, $g$, and the remaining nodes $L = N \setminus \{g\}$ as the lieutenants. The general will broadcast a value to the lieutenants. The goal of the algorithm is for each of the lieutenants to agree on the value sent by $g$ in the presence of a Byzantine faulty general or lieutenants. More precisely, we assume the identity of a sender cannot be spoofed, broadcast communication proceeds in synchronous rounds such that the absence of a message in a round can be detected by the receiver (let $None$ be a constant used when no message is received), and a Byzantine faulty sender can arbitrary corrupt or not send a message, but there are no other faults in the system.

The algorithm proceeds as follows:

\begin{itemize}
\item {\bf $OM(0)$}: $g$ broadcasts a value to each lieutenant, and the lieutenants return the value received (or $None$).
\item {\bf $OM(m)$, $m > 0$}:
  \begin{enumerate}
  \item $g$ broadcasts a value to each lieutenant, $l$.
  \item\label{om:two} Let $v$ be the value received by $l$. Then for each $l$, execute $OM(m-1)$ with general $l$ and lieutenants $L \setminus \{l\}$.
  \item For each lieutenant $l$, output the majority value received from the other lieutenants.
  \end{enumerate}
\end{itemize}

\noindent
Therefore, $OM(1)$ includes two rounds of broadcast communication, one in which the general broadcasts, and one in which the lieutenants exchange their values.

@@

OM(1) is designed to preserve \emph{validity} and \emph{agreement}
properties. Validity states that if the general is non-faulty, then every lieutenant outputs the value sent by the general. More formally, let the general broadcast $v$, and let $l_i$ denote the output for lieutenant $i$:
%
\begin{equation}
  \tag{Validity}
    \forall \,i. \quad l_i = v
\end{equation}
%
Agreement states that each lieutenant outputs the same value. Let $l_i, l_j$ denote the outputs of lieutenants $i, j \in L$, respectively:
%
\begin{equation}
  \tag{Agreement}
    \forall \,i, j. \quad l_i = l_j
\end{equation}
%

$OM(m)$ was originally designed to tolerate Byzantine faults. It is classic
result that $OM(m)$ satisifies validity and agreement if the number of
nodes is at least $3m+1$, and there no more than $m$
Byzantine faulty nodes~\cite{}.

The fault model of $OM(1)$ has been extended to hybrid fault model described in Section~\ref{sec:kibitzer}. Under that fault model, the hypothesis for proving validity and agreement can be weakened to assuming there must be at least $2a+2s+b+1$ nodes, where $a$ is the number of Byzantine (or asymmetric) faults, $s$ is the number of symmetric faults, and $b$ is the number of benign faults. \lee{is the right?}

\subsection{Model Sketch}

@@

@@

@@

@@

>>>>

\lee{note that we use calendar automata for non real-time system}

@@

To demonstrate the model and techniques, we model the Oral Messages Hybrid (OMH) algorithm~\cite{}, a more sophisticated version of the class Oral Messages (OM) algorithm. OM was first formulated by Lamport, Shostak, and Pease~\cite{om} to achieve agreement in a distributed system while tolerating Byzantine faults. OMH was originally developed by Thambidurai and Park~\cite{hybrid} to achieve the same result in the presence of a more sophisticated fault model that includes symmetric and benign faults. However, OMH had a bug, as originally formulated, which was corrected and formally verified by Lincoln and Rushby~\cite{csl-93-2}.

First, we briefly describe the algorithm, sketch our model of the algorithm within the model presented in Section~\ref{sec:model}, then describe it's invariants.

\subsection{OMH(1) Algorithm}
OMH is a recursive algorithm that proceeds in rounds of communication. Let $OMH(m)$ be a recursive function implementing the algorithm, parameterized by the number of rounds, $m$. Consider a finite set of nodes $N$. Distinguish one node as the \emph{general}, $g$, and the remaining nodes $L = N \setminus \{g\}$ as the \emph{lieutenants}. We assume the identity of any general or lieutenant cannot be spoofed. Broadcast communication proceeds in rounds. Denote any message that is detectably faulty (e.g., fails a CRC) or is absent, by $ERR$. Additionally, in the algorithm, nodes report on values they have previously received. In doing so, nodes must differentiate \emph{reporting} $ERR$ from an $ERR$ itself. Let $R$ denote that an error is being reported. Finally, let $V$ be a special, designated value.

%% The goal of the algorithm is for each of the lieutenants to agree on the value sent by $g$ in the presence of a Byzantine faulty general or lieutenants. More precisely, we assume the identity of a sender cannot be spoofed, broadcast communication proceeds in synchronous rounds such that the absence of a message in a round can be detected by the receiver (let $None$ be a constant used when a message is detectably incorrect, including a lack of message), and a Byzantine faulty sender can arbitrary corrupt or not send a message, but there are no other faults in the system.

The algorithm is recursively defined:

\begin{itemize}
\item {\bf $OM(0)$}: $g$ broadcasts a value to each lieutenant, and the lieutenants return the value received (or $ERR$).
\item {\bf $OM(m)$, $m > 0$}:
  \begin{enumerate}
  \item $g$ broadcasts a value to each lieutenant, $l$.
  \item\label{om:two} Let $l_v$ be the value received by $l \in L$ from $g$. Then for each $l$, execute $OMH(m-1)$, assigning $l$ to be the general and $L \setminus \{l\}$ to be the lieutenants. $l$ sends $l_v$ or $R$, if $l_v = ERR$.
  \item For each lieutenant $l \in L$, remove all $ERR$ values received in Step~\ref{om:two} from executing $OHM(m-1)$. Compute the majority value over the remaining values, or $V$ if there is no majority. If the majority value is $R$, return $E$.
  \end{enumerate}
\end{itemize}

\noindent
In particular, $OMH(1)$ includes two rounds of broadcast communication, one in which the general broadcasts, and one in which the lieutenants exchange their values.

@@

OMH is designed to preserve \emph{validity} and \emph{agreement}
properties. Validity states that if the general is non-faulty, then every lieutenant outputs the value sent by the general. More formally, let the general broadcast $v \neq ERR$, and let $l_i$ denote the output for lieutenant $i$:
%
\begin{equation}
  \tag{Validity}
    \forall \,i. \quad l_i = v
\end{equation}
%
Agreement states that each lieutenant outputs the same value. Let $l_i, l_j$ denote the outputs of lieutenants $i, j \in L$, respectively:
%
\begin{equation}
  \tag{Agreement}
    \forall \,i, j. \quad l_i = l_j
\end{equation}
%

We described a hybrid fault model in Section~\ref{sec:kibitzer}. Under that fault model, agreement and validity hold if $2a+2s+b+1 \geq n$, where $n$ is the total number of nodes, $a$ is the number of Byzantine (or asymmetric) faults, $s$ is the number of symmetric faults, and $b$ is the number of benign faults. Additionally, it is required that the number of rounds $m$ be greater or equal to the number of Byzantine faults, $a$~\cite{csl-93-2,hybrid}.

\subsection{Model Sketch}
The verification is carried out in the Symbolic Analysis Laboratory~\cite{}, which contains a suite of model-checkers and a high-level specification language. In our work, we use infinite-state (SMT-based) $k$-induction~\cite{}.

\lee{Maybe talk about MJRTY here. Maybe also talk about relaxing in synchronous communication in model.}

@@

\lee{make clear we're agnostic about human-provided invariants. We conjecture some can be auto-generated. Some might be inferred by IC3.}

@@

Here we present two classes of experimental results. First, we demonstrate the scalability results of the verification, despite the low-level modeling. Then we describe modularity results, demonstrated by making modifications to the model and re-validating the model.

\subsection{Scalability}

@@

\subsection{Modular Verification}

@@

\lee{mention that ASMs similar to disjunction invariants}
